---
title: "R Notebook"
output: html_notebook
---

Question we are trying to answer: what is the average rent for a 1b/1b by district? same for 2b/2b, etc. etc.


#load necessary libraries
```{r}
library(tidyverse)
```
#import/read csv file that our group chose
# i got a message to put the code at the end 


```{r}
df <- read_csv("sf_clean.csv", show_col_types = FALSE)
```

# Calculate average rent grouped by beds

```{r}
avg_rent_beds <- df %>%
  group_by(beds) %>%
  summarize(avg_rent = mean(price, na.rm = TRUE), .groups = "drop")
```


#view results 

```{r}
print(avg_rent_beds)
```
# create a plot: average rent by unit type and  district

```{r}
ggplot(avg_rent_beds, aes(x = factor(beds), y = avg_rent, fill = factor(beds))) +
  geom_col() +
  labs(
    title = "Average Rent by Number of Bedrooms",
    x = "Bedrooms",
    y = "Average Rent (USD)",
    fill = "Bedrooms"
  ) +
  theme_minimal()
```
#Create a comparison table of average rooms and distract

```{r}
avg_rent_table <- df %>%
  group_by(beds, hood_district) %>%
  summarize(avg_rent = mean(price, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(
    names_from = hood_district,
    values_from = avg_rent,
    names_prefix = "District_"
  )

```

#View the results

```{r}
print(avg_rent_table)
```

#Check to see if these prices are ststaically significant, for example with 1bedroom.
#filter 1bedroom results
```{r}
df_1bed <- df %>% filter(beds == 1)

```

# Check distribution of sample sizes by district

```{r}
table(df_1bed$hood_district)
```
# Run ANOVA to compare mean rent across districts
```{r}
anova_result <- aov(price ~ factor(hood_district), data = df_1bed)
```

#View results
```{r}
summary(anova_result)

```

# interpretation
P value is much less than 0.05, meaning. There is a statistically significant difference in average 1-bedroom rent across districts.

