---
title: "R Notebook"
output: html_notebook
---

# Load libraries

```{r}
library(tidyverse)
```

# Load csv

```{r}
sf_rent_df <- read_csv("sf_clean.csv") %>% 
  drop_na() %>% 
print(sf_rent_df)
```

# Create parking and prices data frame

This data frame will not be sorted by number of bedrooms.

```{r}
simplified_df <- sf_rent_df %>% select(price, parking)
print(simplified_df)
```

# Descriptive stats by type of parking

using these descriptive stats, i was able to view mean median and mode rent prices within each parking category

# Create data frames based on price for each parking category

```{r}
valet_df <- simplified_df %>% filter(parking == "(a) valet")
protected_df <- simplified_df %>% filter(parking == "(b) protected")
off_street_df <- simplified_df %>% filter(parking == "(c) off-street")
no_parking_df <- simplified_df %>% filter(parking == "(d) no parking")

summary(valet_df)
summary(protected_df)
summary(off_street_df)
summary(no_parking_df)
```

## Comparing average rent price across the parking categories

```{r}
summary_parking_df <- simplified_df %>%
  group_by(parking) %>%
  summarise(mean_price = mean(price, na.rm = TRUE))
```

## Create a bar graph displaying the average rent prices in each parking category

```{r}
ggplot(summary_parking_df, aes(x = parking, y = mean_price, fill = parking)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("red", "hotpink", "purple", "blue")) +
  labs(title = "Average Price for 1 bed by Parking",
       x = "Parking",
       y = "Average Price") +
  theme_minimal()
```

# Inferential Statistics

## Create linear regression for parking and price

technique: Gaussian linear regression which is non-parametric and does not assume normal distribution null Hypothesis: parking has no effect on price alternative hypothesis: price is affected by parking availability

```{r}
glm(formula = price ~ parking , family = gaussian, data = simplified_df) %>%
  summary()
```

## Conclusion from the Gaussian linear regression: 

If valet parking is the baseline, the price of housing significantly decreases if parking is protected, on the street or not available at all.

```{r}
# Make sure 'parking' is treated as a factor (if it's categorical)
simplified_df <- simplified_df %>%
  mutate(parking = as.factor(parking))

# Fit the model
model <- glm(formula = price ~ parking, family = gaussian, data = simplified_df)

# Get predicted values
simplified_df <- simplified_df %>%
  mutate(predicted_price = predict(model, type = "response"))

# Plot actual data with model predictions
ggplot(simplified_df, aes(x = parking, y = price, fill = parking)) +
  geom_boxplot(alpha = 0.4) +
  stat_summary(aes(y = predicted_price), fun = mean, geom = "point", 
               shape = 21, color = "yellow", size = 3, fill = "yellow") +
  scale_fill_manual(values = c("red", "hotpink", "purple", "blue")) +
  labs(title = "GLM: Price ~ Parking",
       subtitle = "Boxplot of Actual Prices with Predicted Means from GLM",
       x = "Parking Type",
       y = "Price") +
  theme_minimal()
```

```{r}
# Make sure 'parking' is a factor
simplified_df <- simplified_df %>%
  mutate(parking = as.factor(parking))

# Fit the GLM model
model <- glm(price ~ parking, family = gaussian, data = simplified_df)

# Add predicted values to the dataset
simplified_df <- simplified_df %>%
  mutate(predicted_price = predict(model, type = "response"))

# Calculate the mean predicted price for each parking type
mean_predicted_prices <- simplified_df %>%
  group_by(parking) %>%
  summarise(mean_predicted_price = mean(predicted_price))



# Calculate actual mean price for each parking type
actual_means <- simplified_df %>%
  group_by(parking) %>%
  summarise(actual_mean = mean(price))

# Now, make the plot
ggplot(simplified_df, aes(x = parking, y = price, fill = parking)) +
  geom_boxplot(alpha = 0.4) +
  
  # Predicted mean points from the GLM model
  stat_summary(aes(y = predicted_price), fun = mean, geom = "point", 
               shape = 21, color = "yellow", size = 3, fill = "yellow") +
  
  # Actual mean points from the dataset
  geom_point(data = actual_means, aes(x = parking, y = actual_mean), 
             shape = 21, color = "green", size = 3, fill = "green") +
  
  scale_fill_manual(values = c("red", "hotpink", "purple", "blue")) +
  labs(title = "Comparison: Actual vs. Predicted Means for Price by Parking Type",
       subtitle = "Boxplot of Actual Prices with Predicted and Actual Means",
       x = "Parking Type",
       y = "Price") +
  theme_minimal()


print(actual_means)
print(mean_predicted_prices)
```


## Drawing conclusions from the box plot. 
We were able to compare to predicted mean to the actual means and they seem to be in the same position making this model a good fit.  
